What Determines House Prices in Ames, Iowa

### 2022-07-30 Final Draft/paper.tex
Paper Changes

\documentclass{article}
\usepackage{graphicx}
\usepackage{float}
\title{What Determines House Prices in Ames, Iowa}
\author{Bryan Klee and Joe Lucas}
\date{July 31st, 2022}
\begin{document}
	\maketitle
	\section{Abstract}
	Add this at the end

	\section{Introduction}
	Housing is one of the three basic needs of Humans, so house prices can be extremely important to many people. Studies have shown, and a general knowledge of homes would tell you that the most important factors of house price are Location, Quality of Materials, Condition of House and the Age of the home. We set forth to determine whether these really are the most important factors, if there are other factors that play into it as well, or if none of those factors are important at all.

@@

	\subsection{Data}
	We found a data set of house sales in Ames, Iowa between 2006 and 2010 gathered from the Ames Assessor's Office. The data includes almost any kind of attribute a house can have, ranging from how many rooms there are and the square footage of the house, to if there is a pool or garage and what type of road goes to the home. In total there were 2,930 homes sold between 2006 and 2010, each having 79 descriptive attributes. This amount of data is both helpful and hindering. With the large amount of houses and attributes it should give more accurate results, but will require more effort in cleaning the data overall and in narrowing down the data set to only the most important factors.

	\section{Materials and Methods}
	The housing data set we have has 80 variables for each house, which include 23 nominal, 23 ordinal, 14 discrete, and 20 continuous variables. In order to get a more comprehend-able data set we will have to reduce the variables to a set of features we want to look at. In order to do this we create a general category of 4 variables, and these are the four factors we think are most important, Location, Quality, Condition, and Age. Out of these 4 categories we come up with we have 12 features we come down to. These can be just variables in from our raw data or combinations of variables such as Total Square Footage, which is the sum of 1st Floor Living Area Square Footage, 2nd floor and the basement. So we trim the data set of 80 variables down to a feature group of 12 to study.

	Now looking at distribution we can see that some of the features are very skewed one way. We select the features that have a skew value of higher than .5. This ends up being 6? of the features and in order to make them normally distributed and ready for analysis we will log transform them. The following shows the distribution of the features:

	PUT charts of Skew here

@@

	PuT MORE Charts HERE
	Maybe

@@

	Maybe add correlation plot?

	Looking at the categorical we plot the averages against each other to get a visual idea of whats going on. Visually looking at the plots, we see that Neighborhood and Overall Quality look to have statistically significant different means. Since it appears so, we will run an ANOVA test to truly determine if our categorical values are different. After running ANOVA on all our variables we see that the only features that do not have different means are Month Sold and Year Sold. Since they do not have different means we will remove them as they have no impact. Overall Quality, Neighborhoods and Overall Condition do have significantly different means and therefore they may have an impact and we will have to keep them in for analysis.

	Maybe talk about some stuff we did before MVA?

>>>>

\documentclass{article}
\usepackage{graphicx}
\usepackage{subcaption}
\graphicspath{ {./images/} }
\usepackage{float}
\usepackage{multirow}
\title{What Determines House Prices in Ames, Iowa}
\author{Bryan Klee and Joe Lucas}
\date{July 31st, 2022}
\begin{document}
	\maketitle
	\section{Abstract}
	Add this at the end

	\section{Introduction}
	Housing is one of a human's three most basic needs, so home prices will significantly affect everyone's financial health and well-being. Studies have shown, and a general knowledge of homes would tell you, that the most important factors of house price are location, quality of Materials, condition of the home, and the age of the home. We sought to determine whether these really are the most important factors, if any of these factors are more important than anotehr, or if none of those factors are important at all.

@@

	\subsection{Data}
	We found a data set of house sales in Ames, Iowa between 2006 and 2010 gathered from the Ames Assessor's Office. The data includes almost any kind of attribute a house can have, ranging from how many rooms there are and the square footage of the house, to if there is a pool or garage and what type of road goes to the home. In total there were 2,930 homes sold between 2006 and 2010, each having 79 descriptive attributes. This amount of data is both helpful and hindering. With the large amount of houses and attributes it should give more accurate results, but will require more effort in cleaning the data overall and in narrowing down the data set to only the most important factors. We split the data in half, the first half as apart of a training set, and the second half as a validation set.

	\section{Materials and Methods}
	The housing data set we have has 80 variables for each house, which include 23 nominal, 23 ordinal, 14 discrete, and 20 continuous variables. In order to get a more comprehendable data set we will have to reduce the variables to a set of features we want to look at. In order to do this we create a general category of 4 variables, and these are the four factors we think are most important, Location, Quality, Condition, and Age. Out of these 4 categories we come up with we have 12 features we come down to. These can be just variables in from our raw data or combinations of variables such as Total Square Footage, which is the sum of 1st Floor Living Area Square Footage, 2nd floor and the basement. So we trim the data set of 80 variables down to a feature group of 12 to study.

	\subsection{Initial Feature Reduction}

	A number of the features that we decided to keep were an aggregate of separate features that came in the original dataset. Those features are outlined below:

	\begin{enumerate}
		\item Total Square Footage
		\item[\textbullet] Total Basement Square Footage + General Living Area
		\item Total Porch Square Footage
		\item[\textbullet] Open Porch Square Footage + Enclosed Porch Square Footage + Three Screened In Porch Square Footage + Full Screened In Porch Square Footage
		\item Total Bathrooms
		\item[\textbullet] Full Bathrooms + Half Bathrooms
		\item Years Old
		\item[\textbullet] 2022 - Year Built
		\item Lot Frotage
		\item Lot Area
		\item Neighborhood
		\item Overall Quality
		\item Overall Condition
		\item Total Rooms Above Ground
		\item Garage Area
		\item Year Sold
		\item Month Sold
		\item Sale Price
		\item Dependent Variable
	\end{enumerate}

	\subsection{Skewness of Variables}

	Initial Exploration of our features showed that many of them were strongly skewed to the right. A total of six features of the features and in order to make them normally distributed and ready for analysis we will log transform them. A plot of their distributions is shown in the figure below.

	\begin{figure}
		\includegraphics[width=\textwidth]{skewplots}
		\caption{Histograms for the features. Lot Area, Lot Frontage, Total Porch Square Footage, Total Square Footage, Wood Deck Square Footage, and Years Old. Their skew values are 1.54, 12.59, 1.55, 0.65, 2.01, and 0.61, respectively.}
		\label{fig:skew}
	\end{figure}

@@

	%...

\begin{figure}[h!]
	\centering
	\begin{subfigure}[b]{0.4\linewidth}
	  \includegraphics[width=\linewidth]{salehist}
	  \caption{A histogram of the target variable, Sale Price}
	\end{subfigure}
	\begin{subfigure}[b]{0.4\linewidth}
	  \includegraphics[width=\linewidth]{salebox}
	  \caption{A box plot of the target variable, Sale Price. }
	\end{subfigure}
	\caption{Notice the outliers in upper ranges.}
	\label{fig:saleprice}
  \end{figure}

  %...

  \subsection{Correlation of Variables}

@@

	%...

	\begin{figure}[h!]
		\centering
		\begin{subfigure}[b]{0.4\linewidth}
		  \includegraphics[width=\linewidth]{salecorr}
		  \caption{A bar plot of correlation coefficients of numerical variables to the target variable, Sale Price.}
		\end{subfigure}
		\begin{subfigure}[b]{0.4\linewidth}
		  \includegraphics[width=\linewidth]{corrmatrix}
		  \caption{A correlation matrix of all numerical variables in the dataset.}
		\end{subfigure}
		\label{fig:correlation}
	  \end{figure}

	  %...

	\subsection{Analysis of Categorical Variables}

	Looking at the categorical we plot the averages against each other to get a visual idea of whats going on. Visually looking at the plots, we see that Neighborhood and Overall Quality look to have statistically significant different means.

	\begin{figure}
		\includegraphics[width=\textwidth]{catboxes}
		\caption{Box Plots of all categorical variables. Neighborhood, Overall Qual, and Overall Condition have noticably different means.}
		\label{fig:cats}
	\end{figure}

	Since it appears so, we will run an ANOVA test to truly determine if our categorical values are different. After running ANOVA on all our variables we see that the only features that do not have different means are Month Sold and Year Sold. Since they do not have different means we will remove them as they have no impact.

	\begin{table}[]
		\centering
		\begin{tabular}{llll}
		\hline
		\multicolumn{1}{|l|}{Feature} & \multicolumn{1}{l|}{F-Stat} & \multicolumn{1}{l|}{P-value} & \multicolumn{1}{l|}{Significant?} \\ \hline
		Overall Quality               & 374.11                      & 0.00                         & True                              \\
		Neighborhood                  & 74.02                       & 1.63e-230                    & True                              \\
		Overall Condition             & 28.84                       & 1.87e-41                     & True                              \\
		Month Sold                    & 0.98                        & 0.458                        & False                             \\
		Year Sold                     & 0.32                        & 0.865                        & False                            
		\end{tabular}
		\end{table}

	Overall Quality, Neighborhoods and Overall Condition do have significantly different means and therefore they may have an impact and we will have to keep them in for analysis.
